---
title: "Redesigned Visuals"
---

## Data Cleaning

In order to make the original dataset more usable I had to clean it by removing some of the extra rows before reading the file into R. I updated the names for the columns and then I had to pivot the data in order to create the various buckets I needed by murder type.

```{r}

library("tidyverse")
library("tidyr")
library("ggplot2")
library("report")
source("/Users/kwandel/Documents/Personal Stuff/George Mason University/STAT 515 - Spring 2024/R Theme/hw.R")

df <- read.csv(file = "/Users/kwandel/Documents/Personal Stuff/George Mason University/STAT 515 - Spring 2024/Porject Folder/Mid Project/Murrder_By_Weapon.csv")

# Original column names
original_colnames <- colnames(df)

# New column names with spaces instead of periods
new_colnames <- gsub("\\.", " ", original_colnames)

# Rename the columns
colnames(df) <- new_colnames

# Reshape data to long format
df_long <- pivot_longer(df, -Year, names_to = "Offense_Type", values_to = "Count")

```

## Graph 1

The first graph is the original graph that was created but I flipped the y-axis and the corresponding fill to the correct orientation. I also decided to add a few more data points to the graph to make it even easier to understand the change occurring.

```{r}
ggplot(df_long[df_long$Offense_Type == "Firearm", ], aes(x = Year, y = Count)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  geom_area(fill = "lightcoral", alpha = 0.5) +
  geom_text(data = subset(df_long[df_long$Offense_Type == "Firearm", ], Year %% 5 == 0), aes(label = Count), vjust = 1, color = "black", size = 3) +  # Add labels to data points every 5th point
  labs(x = "Year", y = "Murders", title = "Firearm Murders Over Years") +
  hw
```

Fixing the y-axis makes the graph much easier to understand and it can be clearly seen that the number of murders caused by firearms has increased since 2005.

## Graph 2

This next graph is how I would officially redesign the data and would show it in the article.

```{r}
# Filter out "Total Offenses" category
df_filtered <- df_long[df_long$Offense_Type != "Total Offenses", ]

# Plotting
ggplot(df_filtered, aes(x = Year, y = Count, color = Offense_Type)) +
  geom_line() +
  labs(x = "Year", y = "Murders", title = "Murders by Weapon Over Years") +
  hw
```

This graph utilizes the full dataset and looks at the number of murders for each type of offense. By looking at the graph the reader can clearly see that the number of murders using a firearm has increased since 2005. But the reader could also see how many more murders are committed using a firearm compared to other types, which by looking at the graph are decreasing.

## References

```{r}
cite_packages()
```
