---
title: "New Graphs"
---

## Data Cleaning

In order to make the original dataset more usable I had to clean it by removing some of the extra rows before reading the file into R. I updated the names for the columns and then I had to pivot the data in order to create the various buckets I needed by murder type.

```{r}
df <- read.csv(file = "/Users/kwandel/Documents/Personal Stuff/George Mason University/STAT 515 - Spring 2024/Porject Folder/Mid Project/Murrder_By_Weapon.csv")

# Original column names
original_colnames <- colnames(df)

# New column names with spaces instead of periods
new_colnames <- gsub("\\.", " ", original_colnames)

# Rename the columns
colnames(df) <- new_colnames

# Reshape data to long format
df_long <- pivot_longer(df, -Year, names_to = "Offense_Type", values_to = "Count")

```

## Graph 1
The first graph is the original graph that was created but I flipped the y-axis and the corresponding fill to the correct orientation. I also decided to add a few more data points to the graph to make it even easier to understand the change occuring.

```{r}
ggplot(df_long[df_long$Offense_Type == "Firearm", ], aes(x = Year, y = Count)) +
  geom_line(color = "red") +
  geom_point(color = "red") +
  geom_area(fill = "lightcoral", alpha = 0.5) +
  geom_text(data = subset(df_long[df_long$Offense_Type == "Firearm", ], Year %% 5 == 0), aes(label = Count), vjust = 1, color = "black", size = 3) +  # Add labels to data points every 5th point
  labs(x = "Year", y = "Murders", title = "Firearm Murders Over Years") +
  hw
```


## Graph 2
This next graph is how I would officially redesign the data and would show it in the article. 
```{r}
ggplot(df_long, aes(x = Year, y = Count, color = Offense_Type)) +
  geom_line() +
  labs(x = "Year", y = "Murders", title = "Murders by Weapon Over Years") +
  hw
```

This graph utlitizes the full dataset and looks at the number of murders for each type of offense. 

## 
